---
output:
  pdf_document: default
---

# Work with Data in R

Department of Political Science 

Washington University in St. Louis 

<!--
```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = '.')
```
-->

# Today's Goals

The primary goals for this lab will be two fold. First, to become comfortable using basic R data types and commands. Second, to learn to load and inspect data.

# Introduction to R

In this section you will be becoming more familiar with some of the basic data types in R as well as with some of R's built in functions.  If you have any questions remember to ask!

## Object and Variable:

* Assign the value of `10` to `x`: (Hint: you can use `<-` or `=` to assign a variable values)
```{r}
## WRITE ANSWER HERE
x <- 10
```

* Subtract 5 from `x` and then print it out (Hint: you can use the `print()` command)
```{r}
## WRITE ANSWER HERE
x-5
```

* Set `x` to the third power and print it out (Hint: you can do this with `^` and without the `print()` command)
```{r}
## WRITE ANSWER HERE
x^3
```

* Without running the next line, what do you think the result will be? Run the command to see if you were correct. (Hint: R uses P.E.M.D.A.S. when executing calculations)
```{r,eval=FALSE}
2*(300-x)+5
2*(x-100)+2*(x -4)^(1/2)
```

```{r}
## WRITE ANSWER HERE
585
```
\vspace{10pt}

## Vectors:

In this next section, we'll be using a common R data type known as a vector. A vector is a list of items of the same type.  This can be a list of numbers, names, or even dates, but are all the same data type. When we look at columns of a data frame, we are often looking at a vector in R!

* Below is how you declare a vector. Notice how we use `c( )` to denote what items will be in out vector.
```{r}
test_vector <- c(10,9,8,7,6,5,4,3,2,1)
test_vector
```

* We can access items of a vector using `[ ]` after the vector name. The number inside the brackets is the index of the item in the vector. Vector's indices begin at 1.
```{r}
test_vector[2]
```

* Make a vector `y` with all the even numbers between 2 and 20.
```{r}
## WRITE ANSWER HERE
2:20
seq(2, 20, by = 2)
```

* What is the third item in the vector?
```{r}
## WRITE ANSWER HERE
third = seq(2, 20, by = 2)[3]
third
```

* At which index is 14?
```{r}
## WRITE ANSWER HERE
ind = match(14, seq(2, 20, by = 2))
ind
```
\vspace{10pt}


* Using the vector `z` below, find its **length** (the number of items in it), **maximum value**, **minimum value**, **mean value**, and **median value**. (Hint: R has a number of pre-built functions to find this information. Try using functions such as `min()` to find the answers)
\vspace{10pt}
```{r}

z <- sample(100, 15, replace=TRUE)
print(z)
length(z)
max(z)
min(z)
mean(z)
median(z)
```

```{r}
## WRITE ANSWER HERE
```
\vspace{10pt}


# Load and Inspect Data

In this section, we'll learn how to load and inspect data from a source file. The example data is an experiement from Gerber et al (2008). [^1]

## Experiment Overview:
- August 2006 Primary Statewide Election in Michigan.
- Send postcards with randomly assigned message.
    1. No message
    2. Civic Duty Message
    3. You Are Being Studied Message 
    4. Neighborhood Social Pressure Message
- Measured voter turnout for each group.


[^1]: Source: Gerber, Alan S., Donald P. Green, and Christopher W. Larimer. "Social pressure and voter turnout: Evidence from a large-scale field experiment." American political Science review 102.1 (2008): 33-48.


## Load Data
Before loading the data, it is always important to be aware of the working directory!

```{r}
# you should put the whole path below, e.g., read.csv('/Users/yourname/Documents/social.csv')
social <- read.csv("C:/Users/silag/OneDrive/Desktop/Data Science/social.csv") 
```

## Inspect Data
* What's the class of the object `social`? (Hint: You can use the `class()` command)
```{r}
## WRITE ANSWER HERE
class(social)
```

* We can use `head()` to get an idea of how the data is structured.
```{r}
## WRITE ANSWER HERE
head(social)
```

To reference a specific column in our data frame we can use a `$` and the column name. For example `social$message` returns a vector containing the various messages each person received.
```{r}
## TRY IT HERE
social$message
```